<<<<<<< HEAD
---
title: "Hull Course Project Applied Spatial Statistics Fall 2021"
output: html_notebook
---
## set working directory
```{r}
setwd("~/Indiana PhD/IU--Applied Spatial Statistics/Course_Project")
```

## install/load necessary packages
```{r}
# install/load data parsing packages
if(!require(dplyr)){install.packages("dplyr")}
library(dplyr)
if(!require(tidyr)){install.packages("tidyr")}
library(tidyr)
if(!require(data.table)){install.packages("data.table")}
library(data.table)
if(!require(stringr)){install.packages("stringr")}
library(stringr)
if(!require(lubridate)){install.packages("lubridate")}
library(lubridate)
# install/load map and raster packages
if(!require(sf)){install.packages("sf")}
library(sf)
if(!require(tmap)){install.packages("tmap")}
library(tmap)
if(!require(rgdal)){install.packages("rgdal")}
library(rgdal)
if(!require(sp)){install.packages("sp")}
library(sp)
if(!require(raster)){install.packages("raster")}
library(raster)
# install/load exact multinomial test packages
if(!require(EMT)){install.packages("EMT")}
library(EMT)
if(!require(purrr)){install.packages("purrr")}
library(purrr)
```

## import alternative test results
```{r}
aClimateSegmentaClimateSegment <- read.csv("pvaluesaClimateSegmentaClimateSegment.csv", header = T)

bClimateSegmentbClimateSegment <- read.csv("pvaluesbClimateSegmentbClimateSegment.csv", header = T)

cClimateSegmentcClimateSegment <- read.csv("pvaluescClimateSegmentcClimateSegment.csv", header = T)

dClimateSegmentdClimateSegment <- read.csv("pvaluesdClimateSegmentdClimateSegment.csv", header = T)

eClimateSegmenteClimateSegment <- read.csv("pvalueseClimateSegmenteClimateSegment.csv", header = T)

fClimateSegmentfClimateSegment <- read.csv("pvaluesfClimateSegmentfClimateSegment.csv", header = T)

aClimateSegmentbClimateSegment <- read.csv("pvaluesaClimateSegmentbClimateSegment.csv", header = T)

bClimateSegmentcClimateSegment <- read.csv("pvaluesbClimateSegmentcClimateSegment.csv", header = T)

cClimateSegmentdClimateSegment <- read.csv("pvaluescClimateSegmentdClimateSegment.csv", header = T)

dClimateSegmenteClimateSegment <- read.csv("pvaluesdClimateSegmenteClimateSegment.csv", header = T)

eClimateSegmentfClimateSegment <- read.csv("pvalueseClimateSegmentfClimateSegment.csv", header = T)

aClimateSegmentcClimateSegment <- read.csv("pvaluesaClimateSegmentcClimateSegment.csv", header = T)

bClimateSegmentdClimateSegment <- read.csv("pvaluesbClimateSegmentdClimateSegment.csv", header = T)

cClimateSegmenteClimateSegment <- read.csv("pvaluescClimateSegmenteClimateSegment.csv", header = T)

dClimateSegmentfClimateSegment <- read.csv("pvaluesdClimateSegmentfClimateSegment.csv", header = T)

aClimateSegmentdClimateSegment <- read.csv("pvaluesaClimateSegmentdClimateSegment.csv", header = T)

bClimateSegmenteClimateSegment <- read.csv("pvaluesbClimateSegmenteClimateSegment.csv", header = T)

cClimateSegmentfClimateSegment <- read.csv("pvaluescClimateSegmentfClimateSegment.csv", header = T)

aClimateSegmenteClimateSegment <- read.csv("pvaluesaClimateSegmenteClimateSegment.csv", header = T)

bClimateSegmentfClimateSegment <- read.csv("pvaluesbClimateSegmentfClimateSegment.csv", header = T)
```

## append dataframes into one dataframe
```{r}
pvalues1 <- do.call("rbind", list(aClimateSegmentaClimateSegment, bClimateSegmentbClimateSegment, cClimateSegmentcClimateSegment, dClimateSegmentdClimateSegment, 
eClimateSegmenteClimateSegment, 
fClimateSegmentfClimateSegment, 
aClimateSegmentbClimateSegment, 
bClimateSegmentcClimateSegment, 
cClimateSegmentdClimateSegment, 
dClimateSegmenteClimateSegment, 
eClimateSegmentfClimateSegment, 
aClimateSegmentcClimateSegment, 
bClimateSegmentdClimateSegment, 
cClimateSegmenteClimateSegment, 
dClimateSegmentfClimateSegment, 
aClimateSegmentdClimateSegment, 
bClimateSegmenteClimateSegment, 
cClimateSegmentfClimateSegment, 
aClimateSegmenteClimateSegment, 
bClimateSegmentfClimateSegment))
```

## Find best fit model for each species
```{r}
## Find best fit model for each species
pvalues2 <- pvalues1 %>% group_by(speciesName) %>% top_n(1, padj)
## Subset species with a p-value less than 0.05
pvaluesnotsig1 <- pvalues2[pvalues2$padj < 0.05,]
notsigspecies <- as.data.frame(unique(pvaluesnotsig1$speciesName))
colnames(notsigspecies)[1] <- "speciesName"
## Remove species from best fit model that are not significant
altmodelbestfits1 <- pvalues2[!grepl(paste(notsigspecies$speciesName, collapse="|"), pvalues2$speciesName),]
## Subset to keep desired rows
altmodelbestfits <- subset(altmodelbestfits1, select = c(speciesName, padj, model))
numberbestfitspermodel <- as.data.frame(table(altmodelbestfits$model))
## Find highest fit per model
highestfitpermodel <- altmodelbestfits %>% group_by(model) %>% top_n(1, padj)
## export data
write.csv(highestfitpermodel, "highestfitpermodel.csv", row.names = F)
write.csv(altmodelbestfits, "altmodelbestfits.csv", row.names = F)
write.csv(notsigspecies, "altnotsigspecies.csv", row.names = F)
```





=======
---
title: "Hull Course Project Applied Spatial Statistics Fall 2021"
output: html_notebook
---
## set working directory
```{r}
setwd("~/Indiana PhD/IU--Applied Spatial Statistics/Course_Project")
```

## install/load necessary packages
```{r}
# install/load data parsing packages
if(!require(dplyr)){install.packages("dplyr")}
library(dplyr)
if(!require(tidyr)){install.packages("tidyr")}
library(tidyr)
if(!require(data.table)){install.packages("data.table")}
library(data.table)
if(!require(stringr)){install.packages("stringr")}
library(stringr)
if(!require(lubridate)){install.packages("lubridate")}
library(lubridate)
# install/load map and raster packages
if(!require(sf)){install.packages("sf")}
library(sf)
if(!require(tmap)){install.packages("tmap")}
library(tmap)
if(!require(rgdal)){install.packages("rgdal")}
library(rgdal)
if(!require(sp)){install.packages("sp")}
library(sp)
if(!require(raster)){install.packages("raster")}
library(raster)
# install/load exact multinomial test packages
if(!require(EMT)){install.packages("EMT")}
library(EMT)
if(!require(purrr)){install.packages("purrr")}
library(purrr)
```

## import alternative test results
```{r}
aClimateSegmentaClimateSegment <- read.csv("pvaluesaClimateSegmentaClimateSegment.csv", header = T)

bClimateSegmentbClimateSegment <- read.csv("pvaluesbClimateSegmentbClimateSegment.csv", header = T)

cClimateSegmentcClimateSegment <- read.csv("pvaluescClimateSegmentcClimateSegment.csv", header = T)

dClimateSegmentdClimateSegment <- read.csv("pvaluesdClimateSegmentdClimateSegment.csv", header = T)

eClimateSegmenteClimateSegment <- read.csv("pvalueseClimateSegmenteClimateSegment.csv", header = T)

fClimateSegmentfClimateSegment <- read.csv("pvaluesfClimateSegmentfClimateSegment.csv", header = T)

aClimateSegmentbClimateSegment <- read.csv("pvaluesaClimateSegmentbClimateSegment.csv", header = T)

bClimateSegmentcClimateSegment <- read.csv("pvaluesbClimateSegmentcClimateSegment.csv", header = T)

cClimateSegmentdClimateSegment <- read.csv("pvaluescClimateSegmentdClimateSegment.csv", header = T)

dClimateSegmenteClimateSegment <- read.csv("pvaluesdClimateSegmenteClimateSegment.csv", header = T)

eClimateSegmentfClimateSegment <- read.csv("pvalueseClimateSegmentfClimateSegment.csv", header = T)

aClimateSegmentcClimateSegment <- read.csv("pvaluesaClimateSegmentcClimateSegment.csv", header = T)

bClimateSegmentdClimateSegment <- read.csv("pvaluesbClimateSegmentdClimateSegment.csv", header = T)

cClimateSegmenteClimateSegment <- read.csv("pvaluescClimateSegmenteClimateSegment.csv", header = T)

dClimateSegmentfClimateSegment <- read.csv("pvaluesdClimateSegmentfClimateSegment.csv", header = T)

aClimateSegmentdClimateSegment <- read.csv("pvaluesaClimateSegmentdClimateSegment.csv", header = T)

bClimateSegmenteClimateSegment <- read.csv("pvaluesbClimateSegmenteClimateSegment.csv", header = T)

cClimateSegmentfClimateSegment <- read.csv("pvaluescClimateSegmentfClimateSegment.csv", header = T)

aClimateSegmenteClimateSegment <- read.csv("pvaluesaClimateSegmenteClimateSegment.csv", header = T)

bClimateSegmentfClimateSegment <- read.csv("pvaluesbClimateSegmentfClimateSegment.csv", header = T)
```

## append dataframes into one dataframe
```{r}
pvalues1 <- do.call("rbind", list(aClimateSegmentaClimateSegment, bClimateSegmentbClimateSegment, cClimateSegmentcClimateSegment, dClimateSegmentdClimateSegment, 
eClimateSegmenteClimateSegment, 
fClimateSegmentfClimateSegment, 
aClimateSegmentbClimateSegment, 
bClimateSegmentcClimateSegment, 
cClimateSegmentdClimateSegment, 
dClimateSegmenteClimateSegment, 
eClimateSegmentfClimateSegment, 
aClimateSegmentcClimateSegment, 
bClimateSegmentdClimateSegment, 
cClimateSegmenteClimateSegment, 
dClimateSegmentfClimateSegment, 
aClimateSegmentdClimateSegment, 
bClimateSegmenteClimateSegment, 
cClimateSegmentfClimateSegment, 
aClimateSegmenteClimateSegment, 
bClimateSegmentfClimateSegment))
```

## Find best fit model for each species
```{r}
## Find best fit model for each species
pvalues2 <- pvalues1 %>% group_by(speciesName) %>% top_n(1, padj)
## Subset species with a p-value less than 0.05
pvaluesnotsig1 <- pvalues2[pvalues2$padj < 0.05,]
notsigspecies <- as.data.frame(unique(pvaluesnotsig1$speciesName))
colnames(notsigspecies)[1] <- "speciesName"
## Remove species from best fit model that are not significant
altmodelbestfits1 <- pvalues2[!grepl(paste(notsigspecies$speciesName, collapse="|"), pvalues2$speciesName),]
## Subset to keep desired rows
altmodelbestfits <- subset(altmodelbestfits1, select = c(speciesName, padj, model))
numberbestfitspermodel <- as.data.frame(table(altmodelbestfits$model))
## Find highest fit per model
highestfitpermodel <- altmodelbestfits %>% group_by(model) %>% top_n(1, padj)
## export data
write.csv(highestfitpermodel, "highestfitpermodel.csv", row.names = F)
write.csv(altmodelbestfits, "altmodelbestfits.csv", row.names = F)
write.csv(notsigspecies, "altnotsigspecies.csv", row.names = F)
```





>>>>>>> b1fb11f5d93378a74feebd963fd9f1464e8b1d91
