---
title: "Hull Course Project Applied Spatial Statistics Fall 2021"
output: html_notebook
---
## set working directory
```{r}
setwd("~/Indiana PhD/IU--Applied Spatial Statistics/Course_Project")
```

## install/load necessary packages
```{r}
# install/load data parsing packages
if(!require(dplyr)){install.packages("dplyr")}
library(dplyr)
if(!require(tidyr)){install.packages("tidyr")}
library(tidyr)
if(!require(data.table)){install.packages("data.table")}
library(data.table)
if(!require(stringr)){install.packages("stringr")}
library(stringr)
if(!require(lubridate)){install.packages("lubridate")}
library(lubridate)
# install/load map and raster packages
if(!require(sf)){install.packages("sf")}
library(sf)
if(!require(tmap)){install.packages("tmap")}
library(tmap)
if(!require(rgdal)){install.packages("rgdal")}
library(rgdal)
if(!require(sp)){install.packages("sp")}
library(sp)
if(!require(raster)){install.packages("raster")}
library(raster)
# install/load exact multinomial test packages
if(!require(EMT)){install.packages("EMT")}
library(EMT)
if(!require(purrr)){install.packages("purrr")}
library(purrr)
if(!require(reshape2)){install.packages("reshape2")}
library(reshape2)
```

## import data
```{r}
# upload shapefile with Indiana counties
indiana1 <- shapefile("~/Indiana PhD/IU--Applied Spatial Statistics/Course_Project/Indiana_Counties_2/indiana2.shp")
# change CRS to WGS84
indiana <- spTransform(indiana1, CRS=CRS("+proj=longlat +datum=WGS84"))
```

## import nc file containing climate data. Data is originally from: https://www.ncei.noaa.gov/data/nclimgrid-monthly/access/.
```{r}
# import data
climate1.brick <- brick("nclimgrid_tmin.nc")
# transform shapefile CRS to match raster CRS
indiana <- spTransform(indiana, crs(climate1.brick))
```

## subset raster data for study period and calculate a January layer by averaging January values into one raster layer
```{r}
# get layer names
layer_names <- names(climate1.brick)
# create vector of layer names to keep for current time period (January from 2010-2020)
yearsc <- c(2010:2020)
layers_keepc <- paste("X",yearsc,".01",".01",sep="")
# keep layers if they contain current time period
climate2.brickc <- climate1.brick[[layers_keepc]]
```

## mask both rasters by Indiana shapefile
```{r}
climate.brickc = mask(climate2.brickc, indiana)
```

## calculate average minimum January temperature across Indiana per year
```{r}
# calculate average minimum January temperature per year for the current time period
climate.average.yearc <- data.frame(rasterlayer.mean=cellStats(climate.brickc, "mean"))
```

## plot time series of average minimum January temperature across Indiana per year
```{r}
# plot for current time period
plot(yearsc,climate.average.yearc$rasterlayer.mean, type = "l", main = "Average January Minimum Temperature across Indiana", xlab = "Years (2010-2020)", ylab = "Average Minimum Temperature (C)", cex.main=0.9, cex.lab=0.8, xlim=c(2011,2019))
abline(lsfit(yearsc, climate.average.yearc$rasterlayer.mean), col = "red")
year.fitc = lm(climate.average.yearc$rasterlayer.mean~yearsc)
summary(year.fitc)
```

## average minimum average January temperatures per grid across study timeframe
```{r}
# calculate mean per grid and convert to raster for current time period
climate.average.gridc <- mean(climate.brickc)
```

## calculate average minimum January temperature per county
```{r}
# Extract raster values to list object for current time period
grid.raster.valuesc <- raster::extract(climate.average.gridc, indiana)
# Use list apply to calculate mean for each polygon
county.raster.valuesc <- lapply(grid.raster.valuesc, FUN=mean)
# list to dataframe
county.raster.valuesc <- unlist(county.raster.valuesc)
# Join mean values to polygon data
indiana$MEANMINAVC <- county.raster.valuesc
```

# plot mean January minimum temperatures per Indiana county
```{r}
# plot current temperature
tm_shape(indiana, projection="+init=epsg:4326") + tm_polygons("MEANMINAVC", border.col = "grey30", title="Temperature (C)", labels = c("b: -9 to -8", "c: -8 to -7", "d: -7 to -6", "e: -6 to -5", "f: -5 to -4")) + tm_compass(position=c("right", "top"), text.size = .4) + tm_layout(main.title = "Mean Minimum January Temperature (2010-2020)", 
          main.title.size = .8, 
          main.title.fontface = 2, 
          legend.title.fontface = 2,
          legend.title.size = .95,
          title.position = c("center", "top"), 
          legend.text.size=.6, 
          compass.type = "rose",
          inner.margins=c(.1,.1,.2,.4), 
          aes.palette = list(seq = "-RdBu"),
          legend.outside = T, 
          legend.position = c("left", "center")) + tm_scale_bar(size = 0.4, position = c("right", "bottom"))
```

## create list of counties per historical climate segments
```{r}
# generate list of segment numbers
letters <- c("a", "b", "c", "d", "e", "f")
segments <- data.frame(paste(letters, "Climate Segment"))
colnames(segments)[1] <- "Segment"
# add segments column to Indiana polygon data
indiana$HISTORICALSEGMENT <- cut(indiana$MEANMINAV, breaks=c(-4,-5,-6,-7,-8,-9,-10), labels=segments$Segment)
indiana$CURRENTSEGMENT <- cut(indiana$MEANMINAVC, breaks=c(-4,-5,-6,-7,-8,-9,-10), labels=segments$Segment)
indiana$HISTORICALSEGMENT <- as.vector(indiana$HISTORICALSEGMENT)
indiana$CURRENTSEGMENT <- as.vector(indiana$CURRENTSEGMENT)
# create dataframe with county and segments
countysegments <- data.frame(indiana$NAME)
colnames(countysegments)[1] <- "county"
countysegments$historicalsegment <- indiana$HISTORICALSEGMENT
countysegments$currentsegment <- indiana$CURRENTSEGMENT
```

# prepare for multinomial test for segment a
```{r}
# prepare column that defines segments: segment of interest equals true and other equals false
countysegments <- countysegments %>% mutate(test_seg=ifelse(historicalsegment=='a Climate Segment',T,F))
# create odds dataframe
odds <- as.data.frame(unique(countysegments$test_seg))
colnames(odds)[1] <- "Segments"
odds$Odds <- c(0.05,0.95)
# create counts
counts <- as.numeric(rbind(17, 0))
```

# conduct Exact Multinomial Test
```{r}
pvaluesa <- array(c(0), dim = 1)
for(i in 1:1){
  pvaluesa[i] = 
multinomial.test(counts, odds$Odds)$p.value}
```

# prepare for multinomial test for segment b
```{r}
# prepare column that defines segments: segment of interest equals true and other equals false
countysegments <- countysegments %>% mutate(test_seg=ifelse(historicalsegment=='b Climate Segment',T,F))
# count occurrences per true and false segments
counts <- dcast(countysegments, currentsegment ~ test_seg)
# keep only segment of interest
counts <- counts[grepl(paste("b Climate Segment", collapse="|"), counts$currentsegment),]
colnames(counts) <- c("currentsegment", "x", "y")
counts <- as.numeric(rbind(counts$x, counts$y))
```

# conduct Exact Multinomial Test
```{r}
pvaluesb <- array(c(0), dim = 1)
for(i in 1:1){
  pvaluesb[i] = 
multinomial.test(counts, odds$Odds)$p.value}
```

# prepare for multinomial test for segment c
```{r}
# prepare column that defines segments: segment of interest equals true and other equals false
countysegments <- countysegments %>% mutate(test_seg=ifelse(historicalsegment=='c Climate Segment',T,F))
# count occurrences per true and false segments
counts <- dcast(countysegments, currentsegment ~ test_seg)
# keep only segment of interest
counts <- counts[grepl(paste("c Climate Segment", collapse="|"), counts$currentsegment),]
colnames(counts) <- c("currentsegment", "x", "y")
counts <- as.numeric(rbind(counts$x, counts$y))
```

# conduct Exact Multinomial Test
```{r}
pvaluesc <- array(c(0), dim = 1)
for(i in 1:1){
  pvaluesc[i] = 
multinomial.test(counts, odds$Odds)$p.value}
```

# prepare for multinomial test for segment d
```{r}
# prepare column that defines segments: segment of interest equals true and other equals false
countysegments <- countysegments %>% mutate(test_seg=ifelse(historicalsegment=='d Climate Segment',T,F))
# count occurrences per true and false segments
counts <- dcast(countysegments, currentsegment ~ test_seg)
# keep only segment of interest
counts <- counts[grepl(paste("d Climate Segment", collapse="|"), counts$currentsegment),]
colnames(counts) <- c("currentsegment", "x", "y")
counts <- as.numeric(rbind(counts$x, counts$y))
```

# conduct Exact Multinomial Test
```{r}
pvaluesd <- array(c(0), dim = 1)
for(i in 1:1){
  pvaluesd[i] = 
multinomial.test(counts, odds$Odds)$p.value}
```

# prepare for multinomial test for segment e
```{r}
# prepare column that defines segments: segment of interest equals true and other equals false
countysegments <- countysegments %>% mutate(test_seg=ifelse(historicalsegment=='e Climate Segment',T,F))
# count occurrences per true and false segments
counts <- dcast(countysegments, currentsegment ~ test_seg)
# keep only segment of interest
counts <- counts[grepl(paste("e Climate Segment", collapse="|"), counts$currentsegment),]
colnames(counts) <- c("currentsegment", "x", "y")
counts <- as.numeric(rbind(counts$x, counts$y))
```

# conduct Exact Multinomial Test
```{r}
pvaluese <- array(c(0), dim = 1)
for(i in 1:1){
  pvaluese[i] = 
multinomial.test(counts, odds$Odds)$p.value}
```

# prepare for multinomial test for segment f
```{r}
# prepare column that defines segments: segment of interest equals true and other equals false
countysegments <- countysegments %>% mutate(test_seg=ifelse(historicalsegment=='f Climate Segment',T,F))
# count occurrences per true and false segments
counts <- dcast(countysegments, currentsegment ~ test_seg)
# keep only segment of interest
counts <- counts[grepl(paste("f Climate Segment", collapse="|"), counts$currentsegment),]
colnames(counts) <- c("currentsegment", "x", "y")
counts <- as.numeric(rbind(counts$x, counts$y))
```

# conduct Exact Multinomial Test
```{r}
pvaluesf <- array(c(0), dim = 1)
for(i in 1:1){
  pvaluesf[i] = 
multinomial.test(counts, odds$Odds)$p.value}
```

# combine p-values into one dataframe and utilize FDR
```{r}
# combine
pvaluessegments <- as.data.frame(rbind(pvaluesa, pvaluesb, pvaluesc, pvaluesd, pvaluese, pvaluesf))
colnames(pvaluessegments)[1] <- "pvalues"
# FDR
pvaluessegments$padj = p.adjust(pvaluessegments$pvalues, method="fdr")
```

# export indiana shapefile
```{r}
writeOGR(obj=indiana, dsn="Indiana_Counties_3", layer="indiana3", driver="ESRI Shapefile")
```

