<<<<<<< HEAD
---
title: "Hull Course Project Applied Spatial Statistics Fall 2021"
output: html_notebook
---
## set working directory
```{r}
setwd("~/Indiana PhD/IU--Applied Spatial Statistics/Course_Project")
```

## install/load necessary packages
```{r}
# install/load data parsing packages
if(!require(dplyr)){install.packages("dplyr")}
library(dplyr)
if(!require(tidyr)){install.packages("tidyr")}
library(tidyr)
if(!require(data.table)){install.packages("data.table")}
library(data.table)
if(!require(stringr)){install.packages("stringr")}
library(stringr)
if(!require(lubridate)){install.packages("lubridate")}
library(lubridate)
# install/load map and raster packages
if(!require(sf)){install.packages("sf")}
library(sf)
if(!require(tmap)){install.packages("tmap")}
library(tmap)
if(!require(rgdal)){install.packages("rgdal")}
library(rgdal)
if(!require(sp)){install.packages("sp")}
library(sp)
if(!require(raster)){install.packages("raster")}
library(raster)
# install/load exact multinomial test packages
if(!require(EMT)){install.packages("EMT")}
library(EMT)
if(!require(purrr)){install.packages("purrr")}
library(purrr)
if(!require(reshape2)){install.packages("reshape2")}
library(reshape2)
```

## import shapefile data
```{r}
# upload shapefile with Indiana counties
indiana1 <- shapefile("~/Indiana PhD/IU--Applied Spatial Statistics/Course_Project/Indiana_Counties_3/indiana3.shp")
# change CRS to WGS84
indiana <- spTransform(indiana1, CRS=CRS("+proj=longlat +datum=WGS84"))
```

## import climate segment probabilities and site species matrix
```{r}
climatesegmentprobs <- read.csv("climatesegmentprobs.csv", header = T)
sitesp2 <- read.csv("sitesp2.csv")
```

## import historical plant distribution data and subset Passiflora incarnata data
```{r}
sigmodela5 <- read.csv("sigmodela5.csv", header = T)
histincarnata <- sigmodela5[grepl(paste("Passiflora incarnata", collapse="|"), sigmodela5$speciesName),]
```

## generate current data for Passiflora incarnata from data collected past 1952, available at: https://midwestherbaria.org/portal/collections/list.php.
```{r}
# add column of occurrences
incarnata <- as.data.frame(indiana$NAME)
colnames(incarnata)[1] <- "name"
incarnata$occurrences <- 0
# add one for counties with current occurrences of Passiflora incarnata on BONAP: Clark, Floyd, Monroe, Vanderburgh, and Washington
countyrecords <- c("Clark", "Floyd", "Monroe", "Vanderburgh", "Washington")
z <- incarnata[match(countyrecords, incarnata$name), ]
z$occurrences <- 1
incarnata[match(z$name, incarnata$name), ] <- z
```

## generate data to visualize historical and current Passiflora incarnata distributions
```{r}
# subset site species matrix for historical data
passifloraincarnata <- sitesp2[grepl(paste("Passiflora incarnata", collapse="|"), sitesp2$speciesName),]
# keep desired columns
passifloraincarnata <- subset(passifloraincarnata, select = c(county, speciesName, n))
# extract county names from Indiana shapefile
countynames <- as.data.frame(indiana$NAME)
colnames(countynames)[1] <- "county"
# complete county names and fill n2 with zeros for new counties
passifloraincarnata <- complete(passifloraincarnata, countynames, speciesName, fill = list(count = 0))
passifloraincarnata[is.na(passifloraincarnata)] <- 0
# reorder based on Indiana shapefile
passifloraincarnata <- passifloraincarnata[order(match(passifloraincarnata$county,countynames$county)),]
indiana$N1 <- passifloraincarnata$n
# add current data to map
indiana$N2 <- incarnata$occurrences
```

## visualize historical and current Passiflora incarnata distributions
```{r}
sp1 <- tm_shape(indiana, projection="+init=epsg:4326") + tm_polygons("N1", border.col = "grey30", title="Timeframe:\n\nHistoric-\n1896-1952\n\nOccurrences", breaks = c(0,1,2), labels = c("0", "1")) + tm_compass(position=c("right", "top"), text.size = .4) + tm_layout(main.title = "Passiflora incarnata",
          main.title.size = .8,
          main.title.fontface = 3,
          legend.title.fontface = 2,
          legend.title.size = .85,
          title.position = c("center", "top"), 
          compass.type = "rose",
          legend.text.size= .7,
          inner.margins=c(.1,.1,.2,.4), 
          aes.palette = list(seq = "Greens"),
          legend.outside = T, 
          legend.position = c("RIGHT", "top")) + tm_scale_bar(size = 0.4, position = c("right", "BOTTOM"))  

# map historical climate segments
sp2 <- tm_shape(indiana, projection="+init=epsg:4326") + tm_polygons("N2", border.col = "grey30", title="Timeframe:\n\nCurrent-\n2010-2020\n\nOccurrences", breaks = c(0,1,2), labels = c("0", "1")) + tm_compass(position=c("right", "top"), text.size = .4) + tm_layout(main.title = "Passiflora incarnata", 
          main.title.size = .8, 
          main.title.fontface = 3, 
          legend.title.fontface = 2,
          legend.title.size = .85,
          title.position = c("center", "top"), 
          compass.type = "rose",
          legend.text.size=.7,
          inner.margins=c(.1,.1,.2,.4), 
          aes.palette = list(seq = "Greens"),
          legend.outside = T, 
          legend.position = c("RIGHT", "top")) + tm_scale_bar(size = 0.4, position = c("right", "BOTTOM")) 

tmap_arrange(sp1, sp2, ncol = 2, asp = NA)
```

## calculate current segment probabilities
```{r}
# calculate un-weighted chance of occurrence in each county
n <- length(indiana$ALAND)
indiana$UNWCHANCES <- 1/n
# calculate weights based off of county land area
indiana$WEIGHTS <- indiana$ALAND / mean(indiana$ALAND)
# calculate weighted chance of occurrence in one county using area weights
indiana$WEIGHTEDCHANCE <- indiana$UNWCHANCES * indiana$WEIGHTS
```

# calculate the probability an occurrence happens in a current climate segment
```{r}
# sum county occurrences by climate segment to get the probabilities of an occurrence happening in any current climate segment
climatesegmentprobsc <- aggregate(WEIGHTEDCHANCE ~ CURRENT, indiana, sum)
colnames(climatesegmentprobsc)[1] <- "SEGMENT"
name2 <- as.data.frame(indiana$NAME)
name2$SEGMENT <- indiana$CURRENT
name2 <- left_join(name2, climatesegmentprobsc)
indiana$CSCHANCEC <- name2$WEIGHTEDCHANCE
```

## calculate climate segment probabilities based on historical distribution
```{r}
# calculate frequency
histincarnata$frequency <- histincarnata$count/sum(histincarnata$count)
# add null expectations
histincarnata$null <- climatesegmentprobs$WEIGHTEDCHANCE
# calculate weights
meanprob <- mean(histincarnata$null)
histincarnata$weights <- histincarnata$null - meanprob
# adjust frequency so it equals un-weighted probabilities
n <- as.numeric(length(histincarnata$X))
zeros <- as.numeric(colSums(histincarnata[5] == 0))
numbnotzeros <- as.numeric(colSums(histincarnata[5] != 0))
subtract <- zeros * 0.09 / numbnotzeros

for(i in 1:n){
  if(histincarnata$frequency[i]== 0){
    histincarnata$unw_probs[i] <- 0 + 0.09
 } else if(histincarnata$frequency[i] != 0){
    histincarnata$unw_probs[i] <- histincarnata$frequency[i] - subtract
  }
}

## apply weights
histincarnata$weighted_probs <- histincarnata$unw_probs + histincarnata$weights
```

## prepare current distribution data for exact multinomial test using historic climate segments
```{r}
# add historic climate segments
incarnata$historic_seg <- indiana$HISTORI
# aggregate counts by climate segment
incarnata2 <- aggregate(incarnata$occurrences, by=list(historic_seg=incarnata$historic_seg), FUN=sum)
colnames(incarnata2)[2] <- "occurrences"
# add species name column
incarnata2$speciesName <- "Passiflora incarnata"
# create vector with occurrences
incarnata3 <- as.numeric(incarnata2$occurrences)
```

## create array for p-values and conduct exact multinomial test
```{r}
## create empty vector for p-values
pvalues1 <- array(c(0), dim = 1)
## conduct exact multinomial test and output p-values for each species
for(i in 1:1){
  pvalues1[i] = 
multinomial.test(incarnata3, histincarnata$weighted_probs)$p.value}
```

## test against current null model
```{r}
# add historic climate segments
incarnata$current_seg <- indiana$CURRENT
# aggregate counts by climate segment
incarnatacurrent2 <- aggregate(incarnata$occurrences, by=list(current_seg=incarnata$current_seg), FUN=sum)
colnames(incarnatacurrent2)[2] <- "occurrences"
# add species name column
incarnatacurrent2$speciesName <- "Passiflora incarnata"
# create vector with occurrences
incarnatacurrent3 <- as.numeric(incarnatacurrent2$occurrences)
```

## create array for p-values and conduct exact multinomial test
```{r}
## create empty vector for p-values
pvaluesnullcurrent <- array(c(0), dim = 1)
## conduct exact multinomial test and output p-values for each species
for(i in 1:1){
  pvaluesnullcurrent[i] = 
multinomial.test(incarnatacurrent3, climatesegmentprobsc$WEIGHTEDCHANCE)$p.value}
```

## create relevant alternative models: def, ef, and f
```{r}
# get current climate segment names
climatesegmentscurrent <- as.data.frame(unique(indiana$CURRENT))
colnames(climatesegmentscurrent)[1] <- "current_seg"
# add odds column and replicate for each alternative model
climatesegmentscurrent$odds <- 0
defodds <- climatesegmentscurrent
efodds <- climatesegmentscurrent
fodds <- climatesegmentscurrent
# modify alternative probabilities based on model
# def odds
remainder <- (1 - 2 * 0.1) / 3
listvalue <- c("d Climate Segment", "e Climate Segment", "f Climate Segment")
notlistvalue <- c("b Climate Segment", "c Climate Segment")
for(i in 1:5){
  if(defodds$current_seg[i] %in% notlistvalue){
    defodds$odds[i] <- 0 + 0.1
 } else if(defodds$current_seg[i] %in% listvalue){
    defodds$odds[i] <- remainder
  }
}
# ef odds
remainder <- (1 - 3 * 0.1) / 2
listvalue <- c("e Climate Segment", "f Climate Segment")
notlistvalue <- c("b Climate Segment", "c Climate Segment", "d Climate Segment")
for(i in 1:5){
  if(efodds$current_seg[i] %in% notlistvalue){
    efodds$odds[i] <- 0 + 0.1
 } else if(efodds$current_seg[i] %in% listvalue){
    efodds$odds[i] <- remainder
  }
}
# f odds
remainder <- 1 - 4 * 0.1
listvalue <- c("f Climate Segment")
notlistvalue <- c("b Climate Segment", "c Climate Segment", "d Climate Segment", "e Climate Segment")
for(i in 1:5){
  if(fodds$current_seg[i] %in% notlistvalue){
    fodds$odds[i] <- 0 + 0.1
 } else if(fodds$current_seg[i] %in% listvalue){
    fodds$odds[i] <- remainder
  }
}
# weight probabilities based off of land area per current climate segment
meanlandareapercurrentseg <- mean(climatesegmentprobsc$WEIGHTEDCHANCE)
climatesegmentprobsc$altmodelweights <- climatesegmentprobsc$WEIGHTEDCHANCE - meanlandareapercurrentseg
# merge weights with alternative models
climatesegmentaltweightsc <- subset(climatesegmentprobsc, select = c(SEGMENT, altmodelweights))
colnames(climatesegmentaltweightsc)[1] <- "current_seg"
defodds <- merge(defodds, climatesegmentaltweightsc)
efodds <- merge(efodds, climatesegmentaltweightsc)
fodds <- merge(fodds, climatesegmentaltweightsc)
# add weights to unweighted probabilites to get weighted probababilities
defodds$weighted_probs <- defodds$odds + defodds$altmodelweights
efodds$weighted_probs <- efodds$odds + efodds$altmodelweights
fodds$weighted_probs <- fodds$odds + fodds$altmodelweights
```

## create array for p-values and conduct exact multinomial test for def model
```{r}
## create empty vector for p-values
pvaluesdefcurrent <- array(c(0), dim = 1)
## conduct exact multinomial test and output p-values for each species
for(i in 1:1){
  pvaluesdefcurrent[i] = 
multinomial.test(incarnatacurrent3, defodds$weighted_probs)$p.value}
```

## create array for p-values and conduct exact multinomial test for ef model
```{r}
## create empty vector for p-values
pvaluesefcurrent <- array(c(0), dim = 1)
## conduct exact multinomial test and output p-values for each species
for(i in 1:1){
  pvaluesefcurrent[i] = 
multinomial.test(incarnatacurrent3, efodds$weighted_probs)$p.value}
```

## create array for p-values and conduct exact multinomial test for f model
```{r}
## create empty vector for p-values
pvaluesfcurrent <- array(c(0), dim = 1)
## conduct exact multinomial test and output p-values for each species
for(i in 1:1){
  pvaluesfcurrent[i] = 
multinomial.test(incarnatacurrent3, fodds$weighted_probs)$p.value}
```

# create pvalue dataframe and export
```{r}
# add model names
pvalues1 <- as.data.frame(pvalues1)
colnames(pvalues1)[1] <- "pvalue"
pvalues1$model <- "historic"
pvaluesnullcurrent <- as.data.frame(pvaluesnullcurrent)
colnames(pvaluesnullcurrent)[1] <- "pvalue"
pvaluesnullcurrent$model <- "nullcurrent"
pvaluesdefcurrent <- as.data.frame(pvaluesdefcurrent)
colnames(pvaluesdefcurrent)[1] <- "pvalue"
pvaluesdefcurrent$model <- "defcurrent"
pvaluesefcurrent <- as.data.frame(pvaluesefcurrent)
colnames(pvaluesefcurrent)[1] <- "pvalue"
pvaluesefcurrent$model <- "efcurrent"
pvaluesfcurrent <- as.data.frame(pvaluesfcurrent)
colnames(pvaluesfcurrent)[1] <- "pvalue"
pvaluesfcurrent$model <- "fcurrent"
# merge data frames
pvaluespassifloraincarnata <- rbind(pvalues1, pvaluesnullcurrent, pvaluesdefcurrent, pvaluesefcurrent, pvaluesfcurrent)
# add species name
pvaluespassifloraincarnata$species_name <- "Passiflora incarnata"
# export
write.csv(pvaluespassifloraincarnata, "Passiflora_incarnata_results.csv", row.names = F)
```


=======
---
title: "Hull Course Project Applied Spatial Statistics Fall 2021"
output: html_notebook
---
## set working directory
```{r}
setwd("~/Indiana PhD/IU--Applied Spatial Statistics/Course_Project")
```

## install/load necessary packages
```{r}
# install/load data parsing packages
if(!require(dplyr)){install.packages("dplyr")}
library(dplyr)
if(!require(tidyr)){install.packages("tidyr")}
library(tidyr)
if(!require(data.table)){install.packages("data.table")}
library(data.table)
if(!require(stringr)){install.packages("stringr")}
library(stringr)
if(!require(lubridate)){install.packages("lubridate")}
library(lubridate)
# install/load map and raster packages
if(!require(sf)){install.packages("sf")}
library(sf)
if(!require(tmap)){install.packages("tmap")}
library(tmap)
if(!require(rgdal)){install.packages("rgdal")}
library(rgdal)
if(!require(sp)){install.packages("sp")}
library(sp)
if(!require(raster)){install.packages("raster")}
library(raster)
# install/load exact multinomial test packages
if(!require(EMT)){install.packages("EMT")}
library(EMT)
if(!require(purrr)){install.packages("purrr")}
library(purrr)
if(!require(reshape2)){install.packages("reshape2")}
library(reshape2)
```

## import shapefile data
```{r}
# upload shapefile with Indiana counties
indiana1 <- shapefile("~/Indiana PhD/IU--Applied Spatial Statistics/Course_Project/Indiana_Counties_3/indiana3.shp")
# change CRS to WGS84
indiana <- spTransform(indiana1, CRS=CRS("+proj=longlat +datum=WGS84"))
```

## import climate segment probabilities and site species matrix
```{r}
climatesegmentprobs <- read.csv("climatesegmentprobs.csv", header = T)
sitesp2 <- read.csv("sitesp2.csv")
```

## import historical plant distribution data and subset Passiflora incarnata data
```{r}
sigmodela5 <- read.csv("sigmodela5.csv", header = T)
histincarnata <- sigmodela5[grepl(paste("Passiflora incarnata", collapse="|"), sigmodela5$speciesName),]
```

## generate current data for Passiflora incarnata from data collected past 1952, available at: https://midwestherbaria.org/portal/collections/list.php.
```{r}
# add column of occurrences
incarnata <- as.data.frame(indiana$NAME)
colnames(incarnata)[1] <- "name"
incarnata$occurrences <- 0
# add one for counties with current occurrences of Passiflora incarnata on BONAP: Clark, Floyd, Monroe, Vanderburgh, and Washington
countyrecords <- c("Clark", "Floyd", "Monroe", "Vanderburgh", "Washington")
z <- incarnata[match(countyrecords, incarnata$name), ]
z$occurrences <- 1
incarnata[match(z$name, incarnata$name), ] <- z
```

## generate data to visualize historical and current Passiflora incarnata distributions
```{r}
# subset site species matrix for historical data
passifloraincarnata <- sitesp2[grepl(paste("Passiflora incarnata", collapse="|"), sitesp2$speciesName),]
# keep desired columns
passifloraincarnata <- subset(passifloraincarnata, select = c(county, speciesName, n))
# extract county names from Indiana shapefile
countynames <- as.data.frame(indiana$NAME)
colnames(countynames)[1] <- "county"
# complete county names and fill n2 with zeros for new counties
passifloraincarnata <- complete(passifloraincarnata, countynames, speciesName, fill = list(count = 0))
passifloraincarnata[is.na(passifloraincarnata)] <- 0
# reorder based on Indiana shapefile
passifloraincarnata <- passifloraincarnata[order(match(passifloraincarnata$county,countynames$county)),]
indiana$N1 <- passifloraincarnata$n
# add current data to map
indiana$N2 <- incarnata$occurrences
```

## visualize historical and current Passiflora incarnata distributions
```{r}
sp1 <- tm_shape(indiana, projection="+init=epsg:4326") + tm_polygons("N1", border.col = "grey30", title="Timeframe:\n\nHistoric-\n1896-1952\n\nOccurrences", breaks = c(0,1,2), labels = c("0", "1")) + tm_compass(position=c("right", "top"), text.size = .4) + tm_layout(main.title = "Passiflora incarnata",
          main.title.size = .8,
          main.title.fontface = 3,
          legend.title.fontface = 2,
          legend.title.size = .85,
          title.position = c("center", "top"), 
          compass.type = "rose",
          legend.text.size= .7,
          inner.margins=c(.1,.1,.2,.4), 
          aes.palette = list(seq = "Greens"),
          legend.outside = T, 
          legend.position = c("RIGHT", "top")) + tm_scale_bar(size = 0.4, position = c("right", "BOTTOM"))  

# map historical climate segments
sp2 <- tm_shape(indiana, projection="+init=epsg:4326") + tm_polygons("N2", border.col = "grey30", title="Timeframe:\n\nCurrent-\n2010-2020\n\nOccurrences", breaks = c(0,1,2), labels = c("0", "1")) + tm_compass(position=c("right", "top"), text.size = .4) + tm_layout(main.title = "Passiflora incarnata", 
          main.title.size = .8, 
          main.title.fontface = 3, 
          legend.title.fontface = 2,
          legend.title.size = .85,
          title.position = c("center", "top"), 
          compass.type = "rose",
          legend.text.size=.7,
          inner.margins=c(.1,.1,.2,.4), 
          aes.palette = list(seq = "Greens"),
          legend.outside = T, 
          legend.position = c("RIGHT", "top")) + tm_scale_bar(size = 0.4, position = c("right", "BOTTOM")) 

tmap_arrange(sp1, sp2, ncol = 2, asp = NA)
```

## calculate current segment probabilities
```{r}
# calculate un-weighted chance of occurrence in each county
n <- length(indiana$ALAND)
indiana$UNWCHANCES <- 1/n
# calculate weights based off of county land area
indiana$WEIGHTS <- indiana$ALAND / mean(indiana$ALAND)
# calculate weighted chance of occurrence in one county using area weights
indiana$WEIGHTEDCHANCE <- indiana$UNWCHANCES * indiana$WEIGHTS
```

# calculate the probability an occurrence happens in a current climate segment
```{r}
# sum county occurrences by climate segment to get the probabilities of an occurrence happening in any current climate segment
climatesegmentprobsc <- aggregate(WEIGHTEDCHANCE ~ CURRENT, indiana, sum)
colnames(climatesegmentprobsc)[1] <- "SEGMENT"
name2 <- as.data.frame(indiana$NAME)
name2$SEGMENT <- indiana$CURRENT
name2 <- left_join(name2, climatesegmentprobsc)
indiana$CSCHANCEC <- name2$WEIGHTEDCHANCE
```

## calculate climate segment probabilities based on historical distribution
```{r}
# calculate frequency
histincarnata$frequency <- histincarnata$count/sum(histincarnata$count)
# add null expectations
histincarnata$null <- climatesegmentprobs$WEIGHTEDCHANCE
# calculate weights
meanprob <- mean(histincarnata$null)
histincarnata$weights <- histincarnata$null - meanprob
# adjust frequency so it equals un-weighted probabilities
n <- as.numeric(length(histincarnata$X))
zeros <- as.numeric(colSums(histincarnata[5] == 0))
numbnotzeros <- as.numeric(colSums(histincarnata[5] != 0))
subtract <- zeros * 0.09 / numbnotzeros

for(i in 1:n){
  if(histincarnata$frequency[i]== 0){
    histincarnata$unw_probs[i] <- 0 + 0.09
 } else if(histincarnata$frequency[i] != 0){
    histincarnata$unw_probs[i] <- histincarnata$frequency[i] - subtract
  }
}

## apply weights
histincarnata$weighted_probs <- histincarnata$unw_probs + histincarnata$weights
```

## prepare current distribution data for exact multinomial test using historic climate segments
```{r}
# add historic climate segments
incarnata$historic_seg <- indiana$HISTORI
# aggregate counts by climate segment
incarnata2 <- aggregate(incarnata$occurrences, by=list(historic_seg=incarnata$historic_seg), FUN=sum)
colnames(incarnata2)[2] <- "occurrences"
# add species name column
incarnata2$speciesName <- "Passiflora incarnata"
# create vector with occurrences
incarnata3 <- as.numeric(incarnata2$occurrences)
```

## create array for p-values and conduct exact multinomial test
```{r}
## create empty vector for p-values
pvalues1 <- array(c(0), dim = 1)
## conduct exact multinomial test and output p-values for each species
for(i in 1:1){
  pvalues1[i] = 
multinomial.test(incarnata3, histincarnata$weighted_probs)$p.value}
```

## test against current null model
```{r}
# add historic climate segments
incarnata$current_seg <- indiana$CURRENT
# aggregate counts by climate segment
incarnatacurrent2 <- aggregate(incarnata$occurrences, by=list(current_seg=incarnata$current_seg), FUN=sum)
colnames(incarnatacurrent2)[2] <- "occurrences"
# add species name column
incarnatacurrent2$speciesName <- "Passiflora incarnata"
# create vector with occurrences
incarnatacurrent3 <- as.numeric(incarnatacurrent2$occurrences)
```

## create array for p-values and conduct exact multinomial test
```{r}
## create empty vector for p-values
pvaluesnullcurrent <- array(c(0), dim = 1)
## conduct exact multinomial test and output p-values for each species
for(i in 1:1){
  pvaluesnullcurrent[i] = 
multinomial.test(incarnatacurrent3, climatesegmentprobsc$WEIGHTEDCHANCE)$p.value}
```

## create relevant alternative models: def, ef, and f
```{r}
# get current climate segment names
climatesegmentscurrent <- as.data.frame(unique(indiana$CURRENT))
colnames(climatesegmentscurrent)[1] <- "current_seg"
# add odds column and replicate for each alternative model
climatesegmentscurrent$odds <- 0
defodds <- climatesegmentscurrent
efodds <- climatesegmentscurrent
fodds <- climatesegmentscurrent
# modify alternative probabilities based on model
# def odds
remainder <- (1 - 2 * 0.1) / 3
listvalue <- c("d Climate Segment", "e Climate Segment", "f Climate Segment")
notlistvalue <- c("b Climate Segment", "c Climate Segment")
for(i in 1:5){
  if(defodds$current_seg[i] %in% notlistvalue){
    defodds$odds[i] <- 0 + 0.1
 } else if(defodds$current_seg[i] %in% listvalue){
    defodds$odds[i] <- remainder
  }
}
# ef odds
remainder <- (1 - 3 * 0.1) / 2
listvalue <- c("e Climate Segment", "f Climate Segment")
notlistvalue <- c("b Climate Segment", "c Climate Segment", "d Climate Segment")
for(i in 1:5){
  if(efodds$current_seg[i] %in% notlistvalue){
    efodds$odds[i] <- 0 + 0.1
 } else if(efodds$current_seg[i] %in% listvalue){
    efodds$odds[i] <- remainder
  }
}
# f odds
remainder <- 1 - 4 * 0.1
listvalue <- c("f Climate Segment")
notlistvalue <- c("b Climate Segment", "c Climate Segment", "d Climate Segment", "e Climate Segment")
for(i in 1:5){
  if(fodds$current_seg[i] %in% notlistvalue){
    fodds$odds[i] <- 0 + 0.1
 } else if(fodds$current_seg[i] %in% listvalue){
    fodds$odds[i] <- remainder
  }
}
# weight probabilities based off of land area per current climate segment
meanlandareapercurrentseg <- mean(climatesegmentprobsc$WEIGHTEDCHANCE)
climatesegmentprobsc$altmodelweights <- climatesegmentprobsc$WEIGHTEDCHANCE - meanlandareapercurrentseg
# merge weights with alternative models
climatesegmentaltweightsc <- subset(climatesegmentprobsc, select = c(SEGMENT, altmodelweights))
colnames(climatesegmentaltweightsc)[1] <- "current_seg"
defodds <- merge(defodds, climatesegmentaltweightsc)
efodds <- merge(efodds, climatesegmentaltweightsc)
fodds <- merge(fodds, climatesegmentaltweightsc)
# add weights to unweighted probabilites to get weighted probababilities
defodds$weighted_probs <- defodds$odds + defodds$altmodelweights
efodds$weighted_probs <- efodds$odds + efodds$altmodelweights
fodds$weighted_probs <- fodds$odds + fodds$altmodelweights
```

## create array for p-values and conduct exact multinomial test for def model
```{r}
## create empty vector for p-values
pvaluesdefcurrent <- array(c(0), dim = 1)
## conduct exact multinomial test and output p-values for each species
for(i in 1:1){
  pvaluesdefcurrent[i] = 
multinomial.test(incarnatacurrent3, defodds$weighted_probs)$p.value}
```

## create array for p-values and conduct exact multinomial test for ef model
```{r}
## create empty vector for p-values
pvaluesefcurrent <- array(c(0), dim = 1)
## conduct exact multinomial test and output p-values for each species
for(i in 1:1){
  pvaluesefcurrent[i] = 
multinomial.test(incarnatacurrent3, efodds$weighted_probs)$p.value}
```

## create array for p-values and conduct exact multinomial test for f model
```{r}
## create empty vector for p-values
pvaluesfcurrent <- array(c(0), dim = 1)
## conduct exact multinomial test and output p-values for each species
for(i in 1:1){
  pvaluesfcurrent[i] = 
multinomial.test(incarnatacurrent3, fodds$weighted_probs)$p.value}
```

# create pvalue dataframe and export
```{r}
# add model names
pvalues1 <- as.data.frame(pvalues1)
colnames(pvalues1)[1] <- "pvalue"
pvalues1$model <- "historic"
pvaluesnullcurrent <- as.data.frame(pvaluesnullcurrent)
colnames(pvaluesnullcurrent)[1] <- "pvalue"
pvaluesnullcurrent$model <- "nullcurrent"
pvaluesdefcurrent <- as.data.frame(pvaluesdefcurrent)
colnames(pvaluesdefcurrent)[1] <- "pvalue"
pvaluesdefcurrent$model <- "defcurrent"
pvaluesefcurrent <- as.data.frame(pvaluesefcurrent)
colnames(pvaluesefcurrent)[1] <- "pvalue"
pvaluesefcurrent$model <- "efcurrent"
pvaluesfcurrent <- as.data.frame(pvaluesfcurrent)
colnames(pvaluesfcurrent)[1] <- "pvalue"
pvaluesfcurrent$model <- "fcurrent"
# merge data frames
pvaluespassifloraincarnata <- rbind(pvalues1, pvaluesnullcurrent, pvaluesdefcurrent, pvaluesefcurrent, pvaluesfcurrent)
# add species name
pvaluespassifloraincarnata$species_name <- "Passiflora incarnata"
# export
write.csv(pvaluespassifloraincarnata, "Passiflora_incarnata_results.csv", row.names = F)
```


>>>>>>> b1fb11f5d93378a74feebd963fd9f1464e8b1d91
