<<<<<<< HEAD
---
title: "Hull Course Project Applied Spatial Statistics Fall 2021"
output: html_notebook
---
## set working directory
```{r}
setwd("~/Indiana PhD/IU--Applied Spatial Statistics/Course_Project")
```

## install/load necessary packages
```{r}
# install/load data parsing packages
if(!require(dplyr)){install.packages("dplyr")}
library(dplyr)
if(!require(tidyr)){install.packages("tidyr")}
library(tidyr)
if(!require(data.table)){install.packages("data.table")}
library(data.table)
if(!require(stringr)){install.packages("stringr")}
library(stringr)
if(!require(lubridate)){install.packages("lubridate")}
library(lubridate)
# install/load map and raster packages
if(!require(sf)){install.packages("sf")}
library(sf)
if(!require(tmap)){install.packages("tmap")}
library(tmap)
if(!require(rgdal)){install.packages("rgdal")}
library(rgdal)
if(!require(sp)){install.packages("sp")}
library(sp)
if(!require(raster)){install.packages("raster")}
library(raster)
# install/load exact multinomial test packages
if(!require(EMT)){install.packages("EMT")}
library(EMT)
if(!require(purrr)){install.packages("purrr")}
library(purrr)
```

### Part I: Parse plant occurrence data and calculate native species richness per county
## load raw plant occurrence data
```{r}
## load raw plant occurrence data
deam1 <- read.csv(file = "Deam_indiana_raw_data.csv", header = T, na.strings=c("","NA"))
```

## Combine genus and specificEpithet and then remove original columns
```{r}
deam1$speciesName <- paste(deam1$genus,deam1$specificEpithet)
deam2 <- subset(deam1, select = c(family, speciesName, recordedBy, eventDate, country, stateProvince, county, decimalLatitude, decimalLongitude))
```

## Remove data with no GPS coordinates or dates
```{r}
deam3 <- deam2 %>% drop_na()
```

## Remove data that are not identified to species (denoted by "NA") and remove hybrids
```{r}
deam4 <- deam3 %>% filter_all(all_vars(!grepl("NA", .)))
deam5 <- deam4 %>% filter_all(all_vars(!grepl("ó", .)))
deam6 <- deam5 %>% filter_all(all_vars(!grepl("×", .)))
```

## Correct county name spelling
```{r}
## Upload county name list
ctynames <- read.csv("ctynames.csv", header = TRUE, fileEncoding="UTF-8-BOM")
# Test for wrong county spellings
x <- deam6$county %in% ctynames$county
y <- data.frame(x)
z <- data.frame(which(grepl("FALSE", y$x)))
## If z contains data fix
## Correct county name spelling
deam6$county[deam6$county == "knox"] <- "Knox"
deam6$county[deam6$county == "gibson"] <- "Gibson"
deam6$county[deam6$county == "cass"] <- "Cass"
deam6$county[deam6$county == "pike"] <- "Pike"
deam6$county[deam6$county == "orange"] <- "Orange"
deam6$county[deam6$county == "grant"] <- "Grant"
deam6$county[deam6$county == "wayne"] <- "Wayne"
deam6$county[deam6$county == "jasper"] <- "Jasper"
deam6$county[deam6$county == "ELkhart"] <- "Elkhart"
deam6$county[deam6$county == "crawford"] <- "Crawford"
deam6$county[deam6$county == "La Porte"] <- "LaPorte"
# Test for wrong county spellings
x <- deam6$county %in% ctynames$county
y <- data.frame(x)
z <- data.frame(which(grepl("FALSE", y$x)))
```

## Correct Indiana name spelling
```{r}
# Correct indiana to Indiana
deam6$stateProvince[deam6$stateProvince == "indiana"] <- "Indiana"
```

## Correct species names against official list at: https://midwestherbaria.org/portal/checklists/checklist.php?clid=3510&pid=124&dynclid=0.
```{r}
## Upload species name list
indspecies <- read.csv("indspecies.csv", header = TRUE, fileEncoding="UTF-8-BOM")

# Test for wrong species spellings
x <- deam6$speciesName %in% indspecies$ScientificName
y <- data.frame(x)
z <- data.frame(which(grepl("FALSE", y$x)))

## Reassign unnaccepted names to their accepted synonyms

deam6$speciesName[deam6$speciesName == "Andropogon gerardii"] <- "Andropogon gerardi"

deam6$speciesName[deam6$speciesName == "Adicea deamii"] <- "Pilea pumila"

deam6$speciesName[deam6$speciesName == "Baptisia lactea"] <- "Baptisia alba"

deam6$speciesName[deam6$speciesName == "Lacinaria scariosa"] <- "Liatris scariosa"

deam6$speciesName[deam6$speciesName == "Carex striatula"] <- "Carex laxiflora"

deam6$speciesName[deam6$speciesName == "Viburnum pubescens"] <- "Viburnum dentatum"

deam6$speciesName[deam6$speciesName == "Celtis pumila"] <- "Celtis occidentalis"

deam6$speciesName[deam6$speciesName == "Vernonia altissima"] <- "Vernonia gigantea"

deam6$speciesName[grepl("Malus io", deam6$speciesName)] <- "Malus ioensis"

deam6$speciesName[deam6$speciesName == "Botrychium obliquum"] <- "Sceptridium dissectum"

deam6$speciesName[deam6$speciesName == "Arabis viridis"] <- "Borodinia missouriensis"

deam6$speciesName[grepl("tes engelmannii", deam6$speciesName)] <- "Isoetes engelmannii"

deam6$speciesName[deam6$speciesName == "Juncus macer"] <- "Juncus anthelatus"

deam6$speciesName[deam6$speciesName == "Solidago deamii"] <- "Solidago simplex"

deam6$speciesName[deam6$speciesName == "Rosa deamii"] <- "Rosa carolina"

deam6$speciesName[deam6$speciesName == "Panicum deamii"] <- "Dichanthelium ovale"

deam6$speciesName[deam6$speciesName == "Piptatherum pungens"] <- "Piptatheropsis pungens"

deam6$speciesName[deam6$speciesName == "Helianthus rigidus"] <- "Helianthus pauciflorus"

deam6$speciesName[deam6$speciesName == "Malus angustifolia"] <- "Malus ioensis"

deam6$speciesName[deam6$speciesName == "Stachys aff. tenuifolia"] <- "Stachys tenuifolia"

deam6$speciesName[deam6$speciesName == "Rubus satis"] <- "Rubus flagellaris"

# Remove remaining species not included on the official IND list

listhybrids <- c('Lacinaria deamii', 'Crataegus incaedua', 'Rubus profusiflorus', 'Tripleurospermum inodorum', 'Bromus briziformis', 'Rubus serratus', 'Ficus carica', 'Argemone mexicana')
deam7 <- deam6[ !grepl(paste(listhybrids, collapse="|"), deam6$speciesName),]

# Check species again
x <- deam7$speciesName %in% indspecies$ScientificName
y <- data.frame(x)
z <- data.frame(which(grepl("FALSE", y$x)))
```

## remove records of non-native species, using the list of non-native Indiana species located at: https://universalfqa.org/view_database/156. The dataframe "deam8" is the same as the data "Deam_Filtered_Data_10_31_2021."
```{r}
# load nonnatives list
nonnatives <- read.csv(file = "nonnatives_indiana.csv", header = T)
# remove species records that are non-native
deam <- deam7[ !grepl(paste(nonnatives$Scientific.Name, collapse="|"), deam7$speciesName),]
write.csv(deam, "Deam_Filtered_Data_10_31_2021.csv", row.names=F)
=======
---
title: "Hull Course Project Applied Spatial Statistics Fall 2021"
output: html_notebook
---
## set working directory
```{r}
setwd("~/Indiana PhD/IU--Applied Spatial Statistics/Course_Project")
```

## install/load necessary packages
```{r}
# install/load data parsing packages
if(!require(dplyr)){install.packages("dplyr")}
library(dplyr)
if(!require(tidyr)){install.packages("tidyr")}
library(tidyr)
if(!require(data.table)){install.packages("data.table")}
library(data.table)
if(!require(stringr)){install.packages("stringr")}
library(stringr)
if(!require(lubridate)){install.packages("lubridate")}
library(lubridate)
# install/load map and raster packages
if(!require(sf)){install.packages("sf")}
library(sf)
if(!require(tmap)){install.packages("tmap")}
library(tmap)
if(!require(rgdal)){install.packages("rgdal")}
library(rgdal)
if(!require(sp)){install.packages("sp")}
library(sp)
if(!require(raster)){install.packages("raster")}
library(raster)
# install/load exact multinomial test packages
if(!require(EMT)){install.packages("EMT")}
library(EMT)
if(!require(purrr)){install.packages("purrr")}
library(purrr)
```

### Part I: Parse plant occurrence data and calculate native species richness per county
## load raw plant occurrence data
```{r}
## load raw plant occurrence data
deam1 <- read.csv(file = "Deam_indiana_raw_data.csv", header = T, na.strings=c("","NA"))
```

## Combine genus and specificEpithet and then remove original columns
```{r}
deam1$speciesName <- paste(deam1$genus,deam1$specificEpithet)
deam2 <- subset(deam1, select = c(family, speciesName, recordedBy, eventDate, country, stateProvince, county, decimalLatitude, decimalLongitude))
```

## Remove data with no GPS coordinates or dates
```{r}
deam3 <- deam2 %>% drop_na()
```

## Remove data that are not identified to species (denoted by "NA") and remove hybrids
```{r}
deam4 <- deam3 %>% filter_all(all_vars(!grepl("NA", .)))
deam5 <- deam4 %>% filter_all(all_vars(!grepl("ó", .)))
deam6 <- deam5 %>% filter_all(all_vars(!grepl("×", .)))
```

## Correct county name spelling
```{r}
## Upload county name list
ctynames <- read.csv("ctynames.csv", header = TRUE, fileEncoding="UTF-8-BOM")
# Test for wrong county spellings
x <- deam6$county %in% ctynames$county
y <- data.frame(x)
z <- data.frame(which(grepl("FALSE", y$x)))
## If z contains data fix
## Correct county name spelling
deam6$county[deam6$county == "knox"] <- "Knox"
deam6$county[deam6$county == "gibson"] <- "Gibson"
deam6$county[deam6$county == "cass"] <- "Cass"
deam6$county[deam6$county == "pike"] <- "Pike"
deam6$county[deam6$county == "orange"] <- "Orange"
deam6$county[deam6$county == "grant"] <- "Grant"
deam6$county[deam6$county == "wayne"] <- "Wayne"
deam6$county[deam6$county == "jasper"] <- "Jasper"
deam6$county[deam6$county == "ELkhart"] <- "Elkhart"
deam6$county[deam6$county == "crawford"] <- "Crawford"
deam6$county[deam6$county == "La Porte"] <- "LaPorte"
# Test for wrong county spellings
x <- deam6$county %in% ctynames$county
y <- data.frame(x)
z <- data.frame(which(grepl("FALSE", y$x)))
```

## Correct Indiana name spelling
```{r}
# Correct indiana to Indiana
deam6$stateProvince[deam6$stateProvince == "indiana"] <- "Indiana"
```

## Correct species names against official list at: https://midwestherbaria.org/portal/checklists/checklist.php?clid=3510&pid=124&dynclid=0.
```{r}
## Upload species name list
indspecies <- read.csv("indspecies.csv", header = TRUE, fileEncoding="UTF-8-BOM")

# Test for wrong species spellings
x <- deam6$speciesName %in% indspecies$ScientificName
y <- data.frame(x)
z <- data.frame(which(grepl("FALSE", y$x)))

## Reassign unnaccepted names to their accepted synonyms

deam6$speciesName[deam6$speciesName == "Andropogon gerardii"] <- "Andropogon gerardi"

deam6$speciesName[deam6$speciesName == "Adicea deamii"] <- "Pilea pumila"

deam6$speciesName[deam6$speciesName == "Baptisia lactea"] <- "Baptisia alba"

deam6$speciesName[deam6$speciesName == "Lacinaria scariosa"] <- "Liatris scariosa"

deam6$speciesName[deam6$speciesName == "Carex striatula"] <- "Carex laxiflora"

deam6$speciesName[deam6$speciesName == "Viburnum pubescens"] <- "Viburnum dentatum"

deam6$speciesName[deam6$speciesName == "Celtis pumila"] <- "Celtis occidentalis"

deam6$speciesName[deam6$speciesName == "Vernonia altissima"] <- "Vernonia gigantea"

deam6$speciesName[grepl("Malus io", deam6$speciesName)] <- "Malus ioensis"

deam6$speciesName[deam6$speciesName == "Botrychium obliquum"] <- "Sceptridium dissectum"

deam6$speciesName[deam6$speciesName == "Arabis viridis"] <- "Borodinia missouriensis"

deam6$speciesName[grepl("tes engelmannii", deam6$speciesName)] <- "Isoetes engelmannii"

deam6$speciesName[deam6$speciesName == "Juncus macer"] <- "Juncus anthelatus"

deam6$speciesName[deam6$speciesName == "Solidago deamii"] <- "Solidago simplex"

deam6$speciesName[deam6$speciesName == "Rosa deamii"] <- "Rosa carolina"

deam6$speciesName[deam6$speciesName == "Panicum deamii"] <- "Dichanthelium ovale"

deam6$speciesName[deam6$speciesName == "Piptatherum pungens"] <- "Piptatheropsis pungens"

deam6$speciesName[deam6$speciesName == "Helianthus rigidus"] <- "Helianthus pauciflorus"

deam6$speciesName[deam6$speciesName == "Malus angustifolia"] <- "Malus ioensis"

deam6$speciesName[deam6$speciesName == "Stachys aff. tenuifolia"] <- "Stachys tenuifolia"

deam6$speciesName[deam6$speciesName == "Rubus satis"] <- "Rubus flagellaris"

# Remove remaining species not included on the official IND list

listhybrids <- c('Lacinaria deamii', 'Crataegus incaedua', 'Rubus profusiflorus', 'Tripleurospermum inodorum', 'Bromus briziformis', 'Rubus serratus', 'Ficus carica', 'Argemone mexicana')
deam7 <- deam6[ !grepl(paste(listhybrids, collapse="|"), deam6$speciesName),]

# Check species again
x <- deam7$speciesName %in% indspecies$ScientificName
y <- data.frame(x)
z <- data.frame(which(grepl("FALSE", y$x)))
```

## remove records of non-native species, using the list of non-native Indiana species located at: https://universalfqa.org/view_database/156. The dataframe "deam8" is the same as the data "Deam_Filtered_Data_10_31_2021."
```{r}
# load nonnatives list
nonnatives <- read.csv(file = "nonnatives_indiana.csv", header = T)
# remove species records that are non-native
deam <- deam7[ !grepl(paste(nonnatives$Scientific.Name, collapse="|"), deam7$speciesName),]
write.csv(deam, "Deam_Filtered_Data_10_31_2021.csv", row.names=F)
>>>>>>> b1fb11f5d93378a74feebd963fd9f1464e8b1d91
```