---
title: "Hull Course Project Applied Spatial Statistics Fall 2021"
output: html_notebook
---
## set working directory
```{r}
setwd("~/Indiana PhD/IU--Applied Spatial Statistics/Course_Project")
```

## install/load necessary packages
```{r}
# install/load data parsing packages
if(!require(dplyr)){install.packages("dplyr")}
library(dplyr)
if(!require(tidyr)){install.packages("tidyr")}
library(tidyr)
if(!require(data.table)){install.packages("data.table")}
library(data.table)
if(!require(stringr)){install.packages("stringr")}
library(stringr)
if(!require(lubridate)){install.packages("lubridate")}
library(lubridate)
# install/load map and raster packages
if(!require(sf)){install.packages("sf")}
library(sf)
if(!require(tmap)){install.packages("tmap")}
library(tmap)
if(!require(rgdal)){install.packages("rgdal")}
library(rgdal)
if(!require(sp)){install.packages("sp")}
library(sp)
if(!require(raster)){install.packages("raster")}
library(raster)
# install/load exact multinomial test packages
if(!require(EMT)){install.packages("EMT")}
library(EMT)
if(!require(purrr)){install.packages("purrr")}
library(purrr)
```

## load data
```{r}
# load nonnatives list
deam <- read.csv(file = "Deam_Filtered_Data_10_31_2021.csv", header = T)
```

## Create site-species matrix for the complete dataset, treating counties as sites
```{r}
# Subset data to only include species name and county
sitesp1 <- subset(deam, select = c(county, speciesName))
# Sum number of occurrences of each species in each county
sitesp2 <- sitesp1 %>% count(county, speciesName)
# Create site x species matrix by reshaping data
sitesp <- pivot_wider(sitesp2, names_from = speciesName, values_from = n,)
# Replace NA's with 0
sitesp[is.na(sitesp)] <- 0
# convert back to data frame and name counties for sites. 
sitesp <- as.data.frame(sitesp)
rownames(sitesp) <- sitesp[,1]
sitesp <- sitesp[,-1]
```

## Calculate alpha diversity of native plant species per Indiana county using site-species matrix
```{r}
# write species richness function with site species matrix as input
S.obs <- function(x = ""){
  rowSums(x > 0) * 1
}
# calculate site species richness (county) using site species matrix of Indiana native species
obssprichwt <- data.frame(S.obs(sitesp))
colnames(obssprichwt)[1] <- "ALPHADIVERSITY"
# Set row header to new column
obssprichwt$county <- rownames(obssprichwt)
```

## Upload shapefile with Indiana counties
```{r warning=FALSE}
# upload shapefile with Indiana counties
indiana1 <- shapefile("~/Indiana PhD/IU--Applied Spatial Statistics/Course_Project/Indiana_Counties/Indiana_Counties.shp")
# change CRS to WGS84
indiana <- spTransform(indiana1, CRS=CRS("+proj=longlat +datum=WGS84"))
```

## Calculate species density per Indiana county land area
```{r}
# merge species richness with polygon data
spdensity <- merge(indiana,
                     obssprichwt,
                     by.x = "NAME",
                     by.y = "county")
# calculate square kilometers per county land area
spdensity$ALAND_KM2 <- spdensity$ALAND/1000000
# calculate ratio of species per square kilometer of land area for each county
spdensity$SPDENSITY_KM2 <- spdensity$ALPHADIVERSITY/spdensity$ALAND_KM2
```

## plot species density per Indiana county
```{r}
# plot using tmap
tm_shape(spdensity, projection="+init=epsg:4326") +
  tm_polygons("SPDENSITY_KM2", border.col = "grey30", title="Species Richness per km^2") + tm_compass(position=c("right", "top"), text.size = .5) + tm_layout(main.title = "Species Density per Indiana County Land Area", 
          main.title.size = .8, 
          main.title.fontface = 2, 
          legend.title.fontface = 2,
          legend.title.size = .6,
          title.position = c("center", "top"), 
          legend.text.size=.6, compass.type = "rose", inner.margins=c(.1,.1,.2,.4), aes.palette = list(seq = "-Spectral"), legend.outside = T, legend.position = c("right", "top")) + tm_scale_bar(size = 0.5)
```

## calculate correlation between county size and species richness
```{r}
# assess normality via histograms
hist(spdensity$ALAND, main = "Land Area per County", xlab = "Land Area (meters squared)", prob = T)
curve(dnorm(x,
            mean=mean(spdensity$ALAND),sd=sd(spdensity$ALAND)),
      col="darkblue", lwd=2, add=TRUE, yaxt="n")
lines(density(spdensity$ALAND), col="red")
hist(spdensity$ALPHADIVERSITY, main = "Species Richness per County", xlab = "Species Richness", prob = T)
curve(dnorm(x,
            mean=mean(spdensity$ALPHADIVERSITY),sd=sd(spdensity$ALPHADIVERSITY)),
      col="darkblue", lwd=2, add=TRUE, yaxt="n")
lines(density(spdensity$ALPHADIVERSITY), col="red")
# assess normality via qq plots
par(mfrow=c(1,2), pty='s', pch=19)
qqnorm(spdensity$ALAND, main = "Land Area per County", pch=19)
qqline(spdensity$ALAND)
qqnorm(spdensity$ALPHADIVERSITY, main = "Species Richness per County", pch=19)
qqline(spdensity$ALPHADIVERSITY)
# log both variables
spdensity$ALANDLOG <- log(spdensity$ALAND)
spdensity$ALPHADIVERSITYLOG <- log(spdensity$ALPHADIVERSITY)
# Reassess normality via qqplots
par(mfrow=c(1,2), pty='s', pch=19)
qqnorm(spdensity$ALANDLOG, main = "Land Area per County, logged", pch=19)
qqline(spdensity$ALANDLOG)
qqnorm(spdensity$ALPHADIVERSITYLOG, main = "Species Richness per County, logged", pch=19)
qqline(spdensity$ALPHADIVERSITYLOG)
# calculate pearson's correlation
rlog <- cor(spdensity$ALANDLOG, spdensity$ALPHADIVERSITYLOG, method = "pearson")
r <- cor(spdensity$ALAND, spdensity$ALPHADIVERSITY, method = "pearson")
```

# export data
```{r}
write.csv(sitesp2, "sitesp2.csv", row.names = F)
```

