<<<<<<< HEAD
---
title: "Hull Course Project Applied Spatial Statistics Fall 2021"
output: html_notebook
---
## set working directory
```{r}
setwd("~/Indiana PhD/IU--Applied Spatial Statistics/Course_Project")
```

## install/load necessary packages
```{r}
# install/load data parsing packages
if(!require(dplyr)){install.packages("dplyr")}
library(dplyr)
if(!require(tidyr)){install.packages("tidyr")}
library(tidyr)
if(!require(data.table)){install.packages("data.table")}
library(data.table)
if(!require(stringr)){install.packages("stringr")}
library(stringr)
if(!require(lubridate)){install.packages("lubridate")}
library(lubridate)
# install/load map and raster packages
if(!require(sf)){install.packages("sf")}
library(sf)
if(!require(tmap)){install.packages("tmap")}
library(tmap)
if(!require(rgdal)){install.packages("rgdal")}
library(rgdal)
if(!require(sp)){install.packages("sp")}
library(sp)
if(!require(raster)){install.packages("raster")}
library(raster)
# install/load exact multinomial test packages
if(!require(EMT)){install.packages("EMT")}
library(EMT)
if(!require(purrr)){install.packages("purrr")}
library(purrr)
```

## import data
```{r}
# upload shapefile with Indiana counties
indiana1 <- shapefile("~/Indiana PhD/IU--Applied Spatial Statistics/Course_Project/Indiana_Counties/Indiana_Counties.shp")
# change CRS to WGS84
indiana <- spTransform(indiana1, CRS=CRS("+proj=longlat +datum=WGS84"))
```

## import nc file containing climate data. Data is originally from: https://www.ncei.noaa.gov/data/nclimgrid-monthly/access/.
```{r}
# import data
climate1.brick <- brick("nclimgrid_tmin.nc")
# transform shapefile CRS to match raster CRS
indiana <- spTransform(indiana, crs(climate1.brick))
```

## subset raster data for study period and calculate a January layer by averaging January values into one raster layer
```{r}
# get layer names
layer_names <- names(climate1.brick)
# create vector of layer names to keep (January from 1896-1952)
years <- c(1896:1952)
layers_keep <- paste("X",years,".01",".01",sep="")
# keep layers if they contain 1896-1952
climate2.brick <- climate1.brick[[layers_keep]]
```

## mask raster data by Indiana shapefile
```{r}
climate.brick = mask(climate2.brick, indiana)
```

## calculate average minimum January temperature across Indiana per year
```{r}
# calculate average minimum January temperature per year
climate.average.year <- data.frame(rasterlayer.mean=cellStats(climate.brick, "mean"))
```

## plot time series of average minimum January temperature across Indiana per year
```{r}
plot(years,climate.average.year$rasterlayer.mean, type = "l", main = "Average January Minimum Temperature across Indiana", xlab = "Years (1896-1952)", ylab = "Average Minimum Temperature (C)", cex.main=0.9, cex.lab=0.8, xlim=c(1898,1950))
abline(lsfit(years, climate.average.year$rasterlayer.mean), col = "red")
year.fit = lm(climate.average.year$rasterlayer.mean~years)
summary(year.fit)
```

## average minimum average January temperatures per grid across study timeframe
```{r}
# calculate mean per grid and convert to raster
climate.average.grid <- mean(climate.brick)
```

## calculate average minimum January temperature per county
```{r}
# Extract raster values to list object
grid.raster.values <- raster::extract(climate.average.grid, indiana)
# Use list apply to calculate mean for each polygon
county.raster.values <- lapply(grid.raster.values, FUN=mean)
# list to dataframe
county.raster.values <- unlist(county.raster.values)
# Join mean values to polygon data
indiana$MEANMINAV <- county.raster.values
```

# plot mean January minimum temperatures per Indiana county
```{r}
# plot
tm_shape(indiana, projection="+init=epsg:4326") + tm_polygons("MEANMINAV", border.col = "grey30", title="Temperature (C)", n = 6, labels = c("Segment a: -10 to -9", "Segment b: -9 to -8", "Segment c: -8 to -7", "Segment d: -7 to -6", "Segment e: -6 to -5", "Segment f: -5 to -4")) + tm_compass(position=c("right", "top"), text.size = .4) + tm_layout(main.title = "Mean Minimum January Temperature (1896-1952)", 
          main.title.size = .85, 
          main.title.fontface = 2, 
          legend.title.fontface = 2,
          legend.title.size = .95,
          title.position = c("center", "top"), 
          legend.text.size=.6, 
          compass.type = "rose",
          inner.margins=c(.1,.1,.2,.4), 
          aes.palette = list(seq = "-RdBu"),
          legend.outside = T, 
          legend.position = c("right", "top")) + tm_scale_bar(size = 0.4)
```

## export shapefile
```{r}
writeOGR(obj=indiana, dsn="Indiana_Counties_2", layer="indiana2", driver="ESRI Shapefile")
```
=======
---
title: "Hull Course Project Applied Spatial Statistics Fall 2021"
output: html_notebook
---
## set working directory
```{r}
setwd("~/Indiana PhD/IU--Applied Spatial Statistics/Course_Project")
```

## install/load necessary packages
```{r}
# install/load data parsing packages
if(!require(dplyr)){install.packages("dplyr")}
library(dplyr)
if(!require(tidyr)){install.packages("tidyr")}
library(tidyr)
if(!require(data.table)){install.packages("data.table")}
library(data.table)
if(!require(stringr)){install.packages("stringr")}
library(stringr)
if(!require(lubridate)){install.packages("lubridate")}
library(lubridate)
# install/load map and raster packages
if(!require(sf)){install.packages("sf")}
library(sf)
if(!require(tmap)){install.packages("tmap")}
library(tmap)
if(!require(rgdal)){install.packages("rgdal")}
library(rgdal)
if(!require(sp)){install.packages("sp")}
library(sp)
if(!require(raster)){install.packages("raster")}
library(raster)
# install/load exact multinomial test packages
if(!require(EMT)){install.packages("EMT")}
library(EMT)
if(!require(purrr)){install.packages("purrr")}
library(purrr)
```

## import data
```{r}
# upload shapefile with Indiana counties
indiana1 <- shapefile("~/Indiana PhD/IU--Applied Spatial Statistics/Course_Project/Indiana_Counties/Indiana_Counties.shp")
# change CRS to WGS84
indiana <- spTransform(indiana1, CRS=CRS("+proj=longlat +datum=WGS84"))
```

## import nc file containing climate data. Data is originally from: https://www.ncei.noaa.gov/data/nclimgrid-monthly/access/.
```{r}
# import data
climate1.brick <- brick("nclimgrid_tmin.nc")
# transform shapefile CRS to match raster CRS
indiana <- spTransform(indiana, crs(climate1.brick))
```

## subset raster data for study period and calculate a January layer by averaging January values into one raster layer
```{r}
# get layer names
layer_names <- names(climate1.brick)
# create vector of layer names to keep (January from 1896-1952)
years <- c(1896:1952)
layers_keep <- paste("X",years,".01",".01",sep="")
# keep layers if they contain 1896-1952
climate2.brick <- climate1.brick[[layers_keep]]
```

## mask raster data by Indiana shapefile
```{r}
climate.brick = mask(climate2.brick, indiana)
```

## calculate average minimum January temperature across Indiana per year
```{r}
# calculate average minimum January temperature per year
climate.average.year <- data.frame(rasterlayer.mean=cellStats(climate.brick, "mean"))
```

## plot time series of average minimum January temperature across Indiana per year
```{r}
plot(years,climate.average.year$rasterlayer.mean, type = "l", main = "Average January Minimum Temperature across Indiana", xlab = "Years (1896-1952)", ylab = "Average Minimum Temperature (C)", cex.main=0.9, cex.lab=0.8, xlim=c(1898,1950))
abline(lsfit(years, climate.average.year$rasterlayer.mean), col = "red")
year.fit = lm(climate.average.year$rasterlayer.mean~years)
summary(year.fit)
```

## average minimum average January temperatures per grid across study timeframe
```{r}
# calculate mean per grid and convert to raster
climate.average.grid <- mean(climate.brick)
```

## calculate average minimum January temperature per county
```{r}
# Extract raster values to list object
grid.raster.values <- raster::extract(climate.average.grid, indiana)
# Use list apply to calculate mean for each polygon
county.raster.values <- lapply(grid.raster.values, FUN=mean)
# list to dataframe
county.raster.values <- unlist(county.raster.values)
# Join mean values to polygon data
indiana$MEANMINAV <- county.raster.values
```

# plot mean January minimum temperatures per Indiana county
```{r}
# plot
tm_shape(indiana, projection="+init=epsg:4326") + tm_polygons("MEANMINAV", border.col = "grey30", title="Temperature (C)", n = 6, labels = c("Segment a: -10 to -9", "Segment b: -9 to -8", "Segment c: -8 to -7", "Segment d: -7 to -6", "Segment e: -6 to -5", "Segment f: -5 to -4")) + tm_compass(position=c("right", "top"), text.size = .4) + tm_layout(main.title = "Mean Minimum January Temperature (1896-1952)", 
          main.title.size = .85, 
          main.title.fontface = 2, 
          legend.title.fontface = 2,
          legend.title.size = .95,
          title.position = c("center", "top"), 
          legend.text.size=.6, 
          compass.type = "rose",
          inner.margins=c(.1,.1,.2,.4), 
          aes.palette = list(seq = "-RdBu"),
          legend.outside = T, 
          legend.position = c("right", "top")) + tm_scale_bar(size = 0.4)
```

## export shapefile
```{r}
writeOGR(obj=indiana, dsn="Indiana_Counties_2", layer="indiana2", driver="ESRI Shapefile")
```
>>>>>>> b1fb11f5d93378a74feebd963fd9f1464e8b1d91
