<<<<<<< HEAD
---
title: "Hull Course Project Applied Spatial Statistics Fall 2021"
output: html_notebook
---
## set working directory
```{r}
setwd("~/Indiana PhD/IU--Applied Spatial Statistics/Course_Project")
```

## install/load necessary packages
```{r}
# install/load data parsing packages
if(!require(dplyr)){install.packages("dplyr")}
library(dplyr)
if(!require(tidyr)){install.packages("tidyr")}
library(tidyr)
if(!require(data.table)){install.packages("data.table")}
library(data.table)
if(!require(stringr)){install.packages("stringr")}
library(stringr)
if(!require(lubridate)){install.packages("lubridate")}
library(lubridate)
# install/load map and raster packages
if(!require(sf)){install.packages("sf")}
library(sf)
if(!require(tmap)){install.packages("tmap")}
library(tmap)
if(!require(rgdal)){install.packages("rgdal")}
library(rgdal)
if(!require(sp)){install.packages("sp")}
library(sp)
if(!require(raster)){install.packages("raster")}
library(raster)
# install/load exact multinomial test packages
if(!require(EMT)){install.packages("EMT")}
library(EMT)
if(!require(purrr)){install.packages("purrr")}
library(purrr)
```

## upload site species matrix
```{r}
sitesp2 <- read.csv("sitesp2.csv")
```


## Upload shapefile with Indiana counties
```{r warning=FALSE}
# upload shapefile with Indiana counties
indiana1 <- shapefile("~/Indiana PhD/IU--Applied Spatial Statistics/Course_Project/Indiana_Counties_2/indiana2.shp")
# change CRS to WGS84
indiana <- spTransform(indiana1, CRS=CRS("+proj=longlat +datum=WGS84"))
```

## create list of counties per segment
```{r}
# generate list of segment numbers
letters <- c("a", "b", "c", "d", "e", "f")
segments <- data.frame(paste(letters, "Climate Segment"))
colnames(segments)[1] <- "Segment"
# add segment column to indiana polygon data
indiana$SEGMENT <- cut(indiana$MEANMINAV, breaks=c(-4,-5,-6,-7,-8,-9,-10), labels=segments$Segment)
indiana$SEGMENT <- as.vector(indiana$SEGMENT)
# create dataframe with county and segment
countysegments <- data.frame(indiana$NAME)
colnames(countysegments)[1] <- "county"
countysegments$segment <- indiana$SEGMENT
```

# calculate the probability an occurrence happens in a county
```{r}
# calculate un-weighted chance of occurrence in each county
n <- length(indiana$ALAND)
indiana$UNWEIGHTEDCHANCE <- 1/n
# calculate weights based off of county land area
indiana$WEIGHTS <- indiana$ALAND / mean(indiana$ALAND)
# calculate weighted chance of occurrence in one county using area weights
indiana$WEIGHTEDCHANCE <- indiana$UNWEIGHTEDCHANCE * indiana$WEIGHTS
```

# calculate the probability an occurrence happens in a climate segment
```{r}
# sum county occurrences by climate segment to get the probabilities of an occurrence happening in any climate segment
climatesegmentprobs <- aggregate(WEIGHTEDCHANCE ~ SEGMENT, indiana, sum)
name <- as.data.frame(indiana$NAME)
name$SEGMENT <- indiana$SEGMENT
name <- left_join(name, climatesegmentprobs)
indiana$CLIMATESEGCHANCE <- name$WEIGHTEDCHANCE
```

## Assign species data to counties and climate segments
```{r}
## Merge copy of ctydata and specdata
modela <- merge(countysegments, sitesp2, all.x = TRUE)
## Create new column with combined latitudinal segments and species
modela$climatesp <- paste(modela$segment, "Divide", modela$speciesName)
```

## Add one
```{r}
## Create new column with occurence values (1)
modela$occval <- 1
```

## Write modela and climate probabilities to csv
```{r}
write.csv(modela5, "modela5.csv", row.names=F)
write.csv(climatesegmentprobs, "climatesegmentprobs.csv", row.names = F)
=======
---
title: "Hull Course Project Applied Spatial Statistics Fall 2021"
output: html_notebook
---
## set working directory
```{r}
setwd("~/Indiana PhD/IU--Applied Spatial Statistics/Course_Project")
```

## install/load necessary packages
```{r}
# install/load data parsing packages
if(!require(dplyr)){install.packages("dplyr")}
library(dplyr)
if(!require(tidyr)){install.packages("tidyr")}
library(tidyr)
if(!require(data.table)){install.packages("data.table")}
library(data.table)
if(!require(stringr)){install.packages("stringr")}
library(stringr)
if(!require(lubridate)){install.packages("lubridate")}
library(lubridate)
# install/load map and raster packages
if(!require(sf)){install.packages("sf")}
library(sf)
if(!require(tmap)){install.packages("tmap")}
library(tmap)
if(!require(rgdal)){install.packages("rgdal")}
library(rgdal)
if(!require(sp)){install.packages("sp")}
library(sp)
if(!require(raster)){install.packages("raster")}
library(raster)
# install/load exact multinomial test packages
if(!require(EMT)){install.packages("EMT")}
library(EMT)
if(!require(purrr)){install.packages("purrr")}
library(purrr)
```

## upload site species matrix
```{r}
sitesp2 <- read.csv("sitesp2.csv")
```


## Upload shapefile with Indiana counties
```{r warning=FALSE}
# upload shapefile with Indiana counties
indiana1 <- shapefile("~/Indiana PhD/IU--Applied Spatial Statistics/Course_Project/Indiana_Counties_2/indiana2.shp")
# change CRS to WGS84
indiana <- spTransform(indiana1, CRS=CRS("+proj=longlat +datum=WGS84"))
```

## create list of counties per segment
```{r}
# generate list of segment numbers
letters <- c("a", "b", "c", "d", "e", "f")
segments <- data.frame(paste(letters, "Climate Segment"))
colnames(segments)[1] <- "Segment"
# add segment column to indiana polygon data
indiana$SEGMENT <- cut(indiana$MEANMINAV, breaks=c(-4,-5,-6,-7,-8,-9,-10), labels=segments$Segment)
indiana$SEGMENT <- as.vector(indiana$SEGMENT)
# create dataframe with county and segment
countysegments <- data.frame(indiana$NAME)
colnames(countysegments)[1] <- "county"
countysegments$segment <- indiana$SEGMENT
```

# calculate the probability an occurrence happens in a county
```{r}
# calculate un-weighted chance of occurrence in each county
n <- length(indiana$ALAND)
indiana$UNWEIGHTEDCHANCE <- 1/n
# calculate weights based off of county land area
indiana$WEIGHTS <- indiana$ALAND / mean(indiana$ALAND)
# calculate weighted chance of occurrence in one county using area weights
indiana$WEIGHTEDCHANCE <- indiana$UNWEIGHTEDCHANCE * indiana$WEIGHTS
```

# calculate the probability an occurrence happens in a climate segment
```{r}
# sum county occurrences by climate segment to get the probabilities of an occurrence happening in any climate segment
climatesegmentprobs <- aggregate(WEIGHTEDCHANCE ~ SEGMENT, indiana, sum)
name <- as.data.frame(indiana$NAME)
name$SEGMENT <- indiana$SEGMENT
name <- left_join(name, climatesegmentprobs)
indiana$CLIMATESEGCHANCE <- name$WEIGHTEDCHANCE
```

## Assign species data to counties and climate segments
```{r}
## Merge copy of ctydata and specdata
modela <- merge(countysegments, sitesp2, all.x = TRUE)
## Create new column with combined latitudinal segments and species
modela$climatesp <- paste(modela$segment, "Divide", modela$speciesName)
```

## Add one
```{r}
## Create new column with occurence values (1)
modela$occval <- 1
```

## Write modela and climate probabilities to csv
```{r}
write.csv(modela5, "modela5.csv", row.names=F)
write.csv(climatesegmentprobs, "climatesegmentprobs.csv", row.names = F)
>>>>>>> b1fb11f5d93378a74feebd963fd9f1464e8b1d91
```